### 2022-09-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_interpreting_probabilities@6cd0d7001a8751ad4718e5d4366c1e20698f6804 🚀

        {\bfseries 1:}\,Community ecologists are increasingly shifting
from a binary thinking of ecological networks (e.g., do species
interact?) to a more probabilistic perspective (e.g., how likely are
species to interact?). Assuredly, the benefits of representing
ecological interactions as probabilistic events are numerous, from a
better assessment of the spatial variation of interactions to an
increase capacity to reconstruct networks from sparse data.\\%
        {\bfseries 2:}\,However, probabilities need to be used with
caution when working with ecological networks. Indeed, depending on the
system at hand and the method used to build probabilistic networks,
probabilities can have different interpretations that imply different
ways to manipulate them. This is rarely discussed in the literature,
thus impeding our ability to use data on probabilistic interactions
appropriately.\\%
        {\bfseries 3:}\,At the core of these differences lie the
distinction between assessing the likelihood that two groups of
individuals \emph{can} interact and the likelihood that they \emph{will}
interact. This impacts the spatial, temporal, and taxonomic scaling of
interaction probabilities, thus further enlightening the need to
properly define them in their ecological context.\\%
        {\bfseries 4:}\,With these challenges in mind, we propose a
general approach to thinking about probabilities in regards to
ecological interactions, and call for better definitions and
conceptualizations of probabilistic ecological networks, both at the
local and regional scales.\\%

@@

\begin{itemize}
\tightlist
\item
  There are different ways to define and interpret interaction
  probabilities depending on the study system and on the method used to
  build probabilistic networks.
\item
  It is important to document and define what we mean by an interaction
  probability because different definitions can have different
  ecological and statistical implications.
\item
  This paper aims to outline some of the challenges in interpreting
  interaction probabilities and propose an approach to thinking about
  them.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  There is a big difference in how we interpret the probability that two
  species \emph{can} interact (metaweb) and the probability that they
  \emph{will} interact (realized networks).
\item
  Interaction probabilities can be used to describe Boolean interactions
  (e.g., the probability that two species interact) and weighted
  interactions (e.g., the probability distribution of the amount of
  energy flow between two species).
\item
  In many studies, it is not obvious if their numbers represent
  interaction scores or probabilities (in the later case, it is rarely
  specified what these probabilities truly represent).
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  It means the probability that two taxa can interact, regardless of
  biological plasticity and environmental variability.
\item
  One observation is enough to set this probability to one.
\item
  Can we turn this into a local network realisation that is also
  probabilistic and intuitive?
\end{itemize}

@@

@@

\begin{itemize}
\tightlist
\item
  Difference of interpretation between probabilities in networks at the
  individual level and at any higher taxonomic level.
\item
  How does the scale up of the nodes from an individual to population to
  taxonomic group group change our interpretation of interaction
  probabilities? How does the aggregation change the interpretation?
\item
  How is it analogous to spatial and temporal scaling? Basically, all
  kinds of scaling are just different ways to aggregate individuals or
  nodes.
\end{itemize}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Poisot2016StrProa}{}%
Poisot, T., Cirtwill, A.R., Cazelles, K., Gravel, D., Fortin, M.-J. \&
Stouffer, D.B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, 7, 303--312.

@@

>>>>

        {\bfseries 1:}\,Community ecologists are increasingly shifting
from a binary thinking of ecological networks (e.g., do species
interact?) to a more probabilistic perspective (e.g., how likely are
species to interact?). Assuredly, the benefits of representing
ecological interactions as probabilistic events are numerous, from a
better assessment of the spatial variation of interactions to an
increase capacity to reconstruct networks from sparse data.\\%
        {\bfseries 2:}\,However, probabilities need to be used with
caution when working with interactions and ecological networks. Indeed,
depending on the system at hand and the method used to build
probabilistic networks, probabilities can have different interpretations
that imply different ways to manipulate them. This is rarely discussed
in the literature, thus impeding our ability to use data on
probabilistic interactions appropriately.\\%
        {\bfseries 3:}\,At the core of these differences lie the
distinction between assessing the likelihood that two groups of
individuals \emph{can} interact and the likelihood that they \emph{will}
interact. This impacts the spatial, temporal, and taxonomic scaling of
interaction probabilities, thus further enlightening the need to
properly define them in their ecological context.\\%
        {\bfseries 4:}\,With these challenges in mind, we propose a
general approach to thinking about probabilities in regards to
ecological interactions, and call for better definitions and
conceptualizations of probabilistic ecological networks, both at the
local and regional scales.\\%

@@

\begin{itemize}
\tightlist
\item
  There are different ways to define and interpret interaction
  probabilities depending on the study system and on the method used to
  build probabilistic networks.
\item
  It is important to document and define what we mean by an interaction
  probability because different definitions can have different
  ecological and statistical implications and final interpretation.
\item
  This paper aims to outline some of the challenges in interpreting
  interaction probabilities and propose an approach to thinking about
  them.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  There is a big difference in how we interpret the probability that two
  species \emph{can} interact (metaweb) and the probability that they
  \emph{will} interact (realized networks).
\item
  Interaction probabilities can be used to describe Boolean interactions
  (e.g., the probability that two species interact) and weighted
  interactions (e.g., the probability distribution of the amount of
  energy flow between two species).
\item
  In many studies, it is not obvious if their numbers represent
  interaction scores or probabilities (in the latter case, it is rarely
  specified what these probabilities truly represent).
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  It means the probability that two taxa can interact, regardless of
  biological plasticity, environmental variability, or co-occurrence.
\item
  One observation is enough to set this probability to one.
\item
  Can we turn this into a local network realisation that is also
  probabilistic and intuitive?
\end{itemize}

@@

\textbf{Papers:} there might be something in these McLeod \emph{et al.}
(2020); McLeod \emph{et al.} (2021); Wood \emph{et al.} (2015)

@@

\begin{itemize}
\tightlist
\item
  Difference of interpretation between probabilities in networks at the
  individual level and at any higher taxonomic level.
\item
  How does the scaling up of the nodes from an individual to population
  to taxonomic group group change our interpretation of interaction
  probabilities? How does the aggregation change the interpretation?
\item
  How is it analogous to spatial and temporal scaling? Basically, all
  kinds of scaling are just different ways to aggregate individuals or
  nodes.
\end{itemize}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
McLeod, A., Leroux, S.J., Gravel, D., Chu, C., Cirtwill, A.R., Fortin,
M.-J., \emph{et al.} (2021). Sampling and asymptotic network properties
of spatial multi-trophic networks. \emph{Oikos}, n/a.

\leavevmode\hypertarget{ref-McLeod2020EffSpe}{}%
McLeod, A.M., Leroux, S.J. \& Chu, C. (2020). Effects of species traits,
motif profiles, and environment on spatial variation in multi-trophic
antagonistic networks. \emph{Ecosphere}, 11, e03018.

\leavevmode\hypertarget{ref-Poisot2016StrProa}{}%
Poisot, T., Cirtwill, A.R., Cazelles, K., Gravel, D., Fortin, M.-J. \&
Stouffer, D.B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, 7, 303--312.

@@

\leavevmode\hypertarget{ref-Wood2015EffSpa}{}%
Wood, S.A., Russell, R., Hanson, D., Williams, R.J. \& Dunne, J.A.
(2015). Effects of spatial scale of sampling on food web structure.
\emph{Ecology and Evolution}, 5, 3769--3782.
